---
title: Phyllotaxis Sprial and Prime Numbers - Experiment
author: Chisato
date: '2018-10-28'
slug: phyllotaxis-sprial-and-prime-numbers-experiment
categories:
  - R
tags:
  - ggplot2
  - art
description: ''
featured: ''
featuredalt: ''
featuredpath: ''
linktitle: ''
---


I recently tried out [Data Camp's](https://www.datacamp.com) project called ["Phyllotaxis: Draw flowers using mathematics"](https://www.datacamp.com/projects/62).  Now I'm hooked on drawing spirals using [golden angle](https://en.wikipedia.org/wiki/Golden_angle). Also the mathematical art on instructor's blog, [Fronkostin](https://fronkonstin.com/category/drawings/) is just amazing! 

Separately on twitter, I've gotten message on art with prime number, and that got me thinking to experiment with prime number & phyllotaxis flowers. I couldn't figure out how I'd generate prime numbers in R, but I came across site where you can [download prime numbers](http://www.naturalnumbers.org/primes.htm).  


```{r setup_flower_function, fig.width=11, message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(patchwork)

## Read first 10000 digits of prime number!
prime <- read_csv(file="http://www.naturalnumbers.org/P-10000.txt", col_names=F)
names(prime) <- c("nth_prime","prime","int")  ## int = interval from previous prime number


## Function to Draw Frlower
my_flower <- function(points=5000,num_colour=9,col_option="magma",angle=pi*(3-sqrt(5)),...){
  flower <- tibble(
    n = c(1:points), ## change number here to use different # of points
    r = sqrt(n),
    is_prime = n %in% prime$prime, #logical
    colour = n%%num_colour, ## 2,3,6,12,18, seems to bring out the sprial pattern
    x = r * cos(angle*n),
    y = r * sin(angle*n)
  )
  
  prime.cnt <- flower %>% filter(is_prime) %>% count()
  angle_deg <- if(angle==pi*(3-sqrt(5))) {"golden angle!(137.51 degree | 2.4 radian)"} else {paste(round(angle*180/pi,2),"degree | ",round(angle,2),"radian")}
  
  ## Drawing Flower (but not using Prime Number)
  flower_plot <-flower %>%
    filter(!is_prime) %>%
    ggplot(aes(x=x, y=y, colour=colour)) +
    geom_point() +
    geom_path(size=0.01) +
    scale_colour_viridis_c(end=0.8, guide="none", option=col_option) +
    coord_fixed() +
    theme_void(base_family="Roboto Condensed") +
    labs(caption=paste(num_colour, "colours used to plot", points-prime.cnt,"dots.\nAngle Used: ", angle_deg), 
         subtitle="Flower Nibbled by Prime Number Bug")
  
  ## Drawing Flower (only using Prime Number)
  flower_prime <-flower %>%
    filter(is_prime) %>%
    ggplot(aes(x=x, y=y, colour=colour)) +
    geom_point() +
    scale_colour_viridis_c(end=0.8, guide="none", option=col_option) +
    coord_fixed() +
    theme_void(base_family="Roboto Condensed") +
    labs(caption=paste("Numbers between 1 and ",points, "have", prime.cnt," Prime Numbers\n"),
         subtitle="Flower made up by Prime Numbers Only")
  
  #You need to Print
  flower_plot + flower_prime
  
}


```

## Experimenting with Different Variables 

I've wrote function to draw flower as above, so I can now experiment by changing below.

* points = Number of points e.g. Up to what number should we use to draw flower? (up to 104729)
* num_colour = Number of colours to use.  When golden angle is used, seems like multiple of 6 makes colours line up? 
* col_option = I can use magma,viridis, plasma, inferno or cividis here
* angle = Angle to use for drawing spirals.  Default is set to golden angle pi*(3-sqrt(5))


```{r Testing_Function, echo=TRUE, fig.width=11, message=FALSE, warning=FALSE, paged.print=FALSE}

my_flower()  ## drawing with all default values e.g. 5000 points, 9 colours, angle= golden angle

my_flower(angle = 2*pi-(pi*(3-sqrt(5))))  ## changing angle 2pi - golden angle so it looks like it's the reverse of the above.

my_flower(angle=pi/sqrt(7))  ##  Just testing out different angles
my_flower(angle=pi/sqrt(pi), num_col=6*3) ## Using 18 different colours, but hard to distinguish the colour...  
my_flower(angle=exp(1), num_col=10) ## exp(1) = Euler's Number as angle. 

my_flower(angle = 0.3, num_col=3) ## All the black arm goes away when you only use prime number

my_flower(num_col=6*3, col_option="plasma") ## plasma colour palette! 
my_flower(points=10000,num_col=18, col_option="viridis")  ## little over crowded...


```

I could play with different angles all day long! So intriguing...  



