---
title: Utilizing K-means to extract colours from your favourite images
author: Chisato
date: '2018-11-12'
slug: utilizing-k-means-to-extract-colours-from-your-favourite-images
categories:
  - R
tags:
  - ggvoronoi
  - treemap
  - imager
  - tidyverse
  - kmeans
  - plotly
description: ''
featured: ''
featuredalt: ''
featuredpath: ''
linktitle: ''
---

I have been playing with package called [imager](https://cran.r-project.org/web/packages/imager/index.html), documentation on this package was extremely helpful! I have read through "[getting started](https://cran.r-project.org/web/packages/imager/vignettes/gettingstarted.html)" as well as [few other tutorial & examples](http://dahtah.github.io/imager/).

I love colours... Tools like [colourlovers](https://www.colourlovers.com/), [Adobe Color CC](https://color.adobe.com/), [Canva Color Palette Generators](https://www.canva.com/color-palette/) are great extracting colours from photo (or in general just getting [inspirations](https://www.canva.com/learn/100-color-combinations/) for colour palette), but I wanted to figure out the way to extract colours from image using R.


## Preps 

Loading up packages & loading up image to R, and extracting RGB info out of image. 
First I need to load up image. I'm using below abstract image with bunch of different colours that I've created with Photoshop just for fun. 

```{r preperation, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse) ##
library(imager) ## such a fun package.  I want to learn more.
library(treemap) ## 
library(ggvoronoi) ## because I'm currently addicted to voronoi diagram.

## Load up the image using load.image function! 
im <- load.image("https://farm4.staticflickr.com/3316/3333507738_9d36d39f6d_b.jpg") ## colourful abstract image. 
#im <- load.image("https://farm9.staticflickr.com/8125/8659010017_54a885f12a_z.jpg") ## mainly blue
#im <- load.image("https://farm2.staticflickr.com/1939/30915465767_2d9a733510_z.jpg") ## animoji! 

## View the image I just loaded. 
plot(im, main="Original Image I Want To Get Some Colours Out Of!")
df_size <-dim(im)[1]*dim(im)[2]  ## numbers of row I'd get if i use it as it is...
max_row_num <- 150000 ## number of maximum row I want (just to limit my size!)

## If df is too big, it's too slow to process on my computer, so shrink the image
shrink_ratio  <- if(df_size > max_row_num) {max_row_num / df_size } else {1}
im <- im %>% imresize(shrink_ratio)
# plot(im) if you want to check how the image has been resized.

## get RGB value at each pixel
im_rgb <- im %>% 
  as.data.frame(wide="c") %>%
  rename(red=c.1,green=c.2,blue=c.3) %>%
  mutate(hexvalue = rgb(red,green,blue)) ## you can create hexvalue using red, green blue value!

## turn image into Grayscale and get luminance "value" too. 
im_gray <- im %>%
  grayscale() %>%
  as.data.frame() 

## combine RGB info and Luminance Value Dataset together.
im_df <- im_rgb %>% 
  inner_join(im_gray) 

im_df %>% head()

```


## Applying K-Means Algorithm on RGB value from Image

Using kmeans function is pretty simple, I'm selecting 12 as k in below example, simply because I wanted to get 12 distinct colours from the picture.  

```{r kmeans_algorithm, echo=TRUE, fig.height=9, fig.width=16, message=FALSE, warning=FALSE, paged.print=FALSE}

## Pick k value to run kMean althorithm.
## But to extract colours, I'd pick k as number I want back! 
my_k <- 12

## Running kmeans algorithm on red, green and blue value to gather similar colour together
kmean_rgb <- kmeans(im_df %>% select(red,green,blue), centers=my_k)

## append cluster id to im_df datasets.
im_df$cluster_num <- kmean_rgb$cluster

## center values can be used as cluster colour! 
kmean_center <- kmean_rgb$centers %>% as.data.frame() %>% 
  mutate(group_hex = rgb(red,green,blue), cluster_num = row_number()) %>%
  inner_join(im_df %>% count(cluster_num))

## I can also save the colour palette for future use as well.
my_colour <- kmean_center$group_hex
my_colour

```

### Viewing Colour Palettes 

I wanted to view the colour palette more visually, instead of HexValues....

```{r colour_palette, echo=TRUE, fig.height=6, fig.width=12, message=FALSE, warning=FALSE, paged.print=FALSE}

num_row = floor(my_k/4)
kmean_center %>% 
  ggplot(aes(x=(cluster_num-1)%%num_row,y=floor((cluster_num-1)/num_row))) + 
  geom_tile(aes(fill=group_hex)) +
  geom_label(aes(label=paste(cluster_num,":",group_hex,"\n",n,"distinct hexvalues")), 
             family="Roboto Condensed", lineheight=0.8) +
  scale_fill_manual(values=sort(kmean_center$group_hex), guide="none") +
  theme_void(base_family="Roboto Condensed") +
  labs(subtitle=paste0("k-Mean clustering center colour with ", my_k, " clusters" )) +
  scale_y_reverse()

```


## Which colours get clustered together?

I wanted to see which colours were bundled together in same cluster. I took a sample because plotting more colours simply just took way too long on my machine...

```{r treemap_clustered_colour, fig.height=9, fig.width=16, message=FALSE, warning=FALSE, paged.print=FALSE}

im_df %>% 
  sample_n(size=10000) %>% ## I'm just going to take sample to make the drawing bit faster...
  count(cluster_num,hexvalue) %>%
  treemap(index=c("cluster_num","hexvalue"),
          type="color",
          vSize="n", 
          vColor="hexvalue",
          algorithm = "squarified",
          fontfamily.labels="Roboto Condensed", 
          fontfamily.title="Roboto Condensed",
          border.col=c("#ffffff","#ffffff50"),
          fontsize.labels=c(24,0),
          aspRatio=16/9,
          title="Clustering with RGB")

```

## Bonus: Voronoi Abstract Art 

Below is more fun to do on other images, but I wanted to just use [ggvoronoi](https://cran.r-project.org/web/packages/ggvoronoi/index.html) package because I'm [currently in love](https://chichacha.netlify.com/2018/11/10/voronoi-diagram-with-ggvoronoi-package-with-train-station-data/)... 

Looks like stained glass! :) 

```{r voronoi_art, fig.height=9, fig.width=16, message=FALSE, warning=FALSE, paged.print=FALSE}

## Create Mini sets, choose to select grayscale value that are bigger more often.
im_df_mini <-im_df %>%
  sample_n(size=1000, weight=(1-value)) 

im_df_mini %>%
  ggplot(aes(x=x,y=y)) +
  geom_voronoi(aes(fill=hexvalue), color="#000000", size=0.1) +
  scale_fill_manual(values=sort(unique(im_df_mini$hexvalue)), guide="none") +
  theme_void() +
  scale_y_reverse() +
  coord_fixed()


```


## Using Plotly package

I've always been curious to use plotly, but I haven't had chance to yet.  Since RGB values are 3 different values, I wanted to use xyz axis to plot (and I didn't know how to do that in ggplot2), so here's yet another way to view how the colour got clustered more visually.

```{r plotly_3d, message=FALSE, warning=FALSE, paged.print=FALSE}

library(plotly)
im_df %>% sample_n(size=2000) %>% 
  plot_ly(x = ~red, y=~blue, z=~green, color=~cluster_num, colors=my_colour, size=~value) %>%
  add_markers()   

```

  


