---
title: Making Calendar with ggplot + Moon Phase Calendar for fun
author: Chisato
date: '2018-05-26'
slug: making-calendar-with-ggplot-moon-phase-calendar
categories:
  - R
tags:
  - ggplot2
  - suncalc
  - lubridate
description: 'Making calendar using ggplot2 + using Moon Phase font'
featured: ''
featuredalt: ''
featuredpath: ''
linktitle: ''
output: 
  html_document: 
    fig_height: 11
    fig_width: 9
---

## Making Calendar with ggplot2

I recently discovered font called [Moon Phase](https://www.dafont.com/moon-phases.font).
I also recently discovered R package, [suncalc](https://CRAN.R-project.org/package=suncalc), and I can get Moon Illumination data by day, so I thought I'd use both to create a simple calendar using [ggplot2](http://ggplot2.tidyverse.org/).



```{r package_setup, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

library(tidyverse) 
library(suncalc)  ## https://CRAN.R-project.org/package=suncalc
library(scales) ## So I can use date_breaks etc on chart
library(lubridate) ## for dealing with time!
library(hrbrthemes)  ## I love this themes for ggplot! 
library(knitr)
library(kableExtra)
library(DT)

```


### Prepping The Data

To make calendar, I need to strip out weekday, month, day, week number within a month. So I can use weekday as x-axis, week number within a month as y-axis, and facet by month. 

```{r calendar_dataprep, echo=TRUE, message=FALSE, warning=FALSE}
## use getMoonIllusmination function to get moon fraction, phase, and angle for 2018.
Moon_2018 <-getMoonIllumination(date = seq.Date(as.Date("2018-01-01"), as.Date("2018-12-31"), by = 1), 
                                 keep = c("fraction", "phase", "angle"))


## Append Extra Info, so I can use ggplot to make calendar. 
Moon_2018_df <- Moon_2018 %>% 
  mutate(wkdy = fct_inorder(weekdays(date, abbreviate=T)), ## 2018 worked well because it starts on Monday
         wkn = week(date), ## Week Number
         mo = month(date, label=T, abbr=T), ## Month
         day = day(date), ## Day
         moon.phase = cut(#phase,breaks=seq(0,1,by=1/28), ## this doesn't work...
                          phase,breaks=seq(0,1,by=1/56),
                          ordered_result=T)) %>%
  group_by(mo) %>% 
  mutate(wkn.mo = dense_rank(wkn)) %>%
  ungroup()

## Simply cutting into 28 blocks didn't work out..  
moon.font.tb <- tibble(
  moon.font = c(NA, letters[rep(1:13,rep(2,13))], 0,0, letters[rep(14:26,rep(2,13))],NA),
  moon.phase = levels(Moon_2018_df$moon.phase)
)

Moon_2018_df <- Moon_2018_df %>% left_join(moon.font.tb)

## Show the data around today to see what data above looks like.
Moon_2018_df %>% filter(between(date,today()-3,today()+3)) %>% kable()

```

## Just Simple Calendar

First I just made simple calendar with below code.

```{r simple_calendar, fig.height=15, fig.width=11, message=FALSE, warning=FALSE, paged.print=FALSE}

##  Here's Just Calendar Part
Moon_2018_df %>% ggplot(aes(x=wkdy, y=wkn.mo)) + 
  geom_point(alpha=0.3, aes(color=wkdy), size=8) + 
  geom_text(aes(label=day),family="Roboto Condensed") +
  facet_wrap(~mo, scales="free_x", ncol=3) +  ## So that Each Line is Quarter!
  scale_x_discrete() +
  scale_y_reverse(breaks=NULL) +
  scale_colour_viridis_d(guide="none") +
  theme_ipsum_rc() +
  labs(title="2018", x="Start of week is Monday", y="") +
  expand_limits(y=c(0.5,6.5)) 


```

## Moon Phase Calendar

This time, I tried using geom_tile function to create tiles.   I've coloured sell using fraction (illuminated fraction of the moon). 0 is New Moon, and 1 is Full Moon.  When it's Full moon sky is lighter, so I've coloured cell little bit lighter. 

```{r moon_phase_calendar, fig.height=15, fig.width=11, message=FALSE, warning=FALSE, paged.print=FALSE}

Moon_2018_df %>% ggplot(aes(x=wkdy, y=wkn.mo)) + 
  geom_tile(alpha=0.8, aes(fill=fraction)) + 
  geom_text(aes(label=paste0(" ",day)),size=3,family="Roboto Condensed",hjust=0, color="white") +
  geom_text(aes(label=moon.font), family="Moon Phases", hjust=1, color="white", size=5) +
  facet_wrap(~mo, scales="free_x", ncol=3) +  ## So that Each Line is Quarter!
  scale_x_discrete() +
  scale_y_reverse(breaks=NULL) +
  scale_fill_viridis_c(name="Moon Illumination", option="inferno",  begin=0, end=0.25, guide="none") +  ## so that sky is darkest when there's new moon
  theme_ipsum_rc() +
  labs(title="2018 Moon Phase Calendar", x="Start of week is Monday", y="") +
  expand_limits(y=c(0.5,6.5)) +
  theme(legend.position="top")

```


## Yet Another Calendar

```{r moon_phase_calender2, fig.height=9, fig.width=11, message=FALSE, warning=FALSE, paged.print=FALSE}

Moon_2018_df %>% 
  ggplot(aes(x=day, y=fct_rev(mo))) + 
  theme_minimal(base_family="Roboto Condensed") + 
  scale_x_continuous(breaks=c(1:31), position="top") +
  geom_text(aes(label=moon.font, color=fraction),family="Moon Phases", hjust=0.5, vjust=0.5, size=8, alpha=0.5) +
  geom_text(aes(label=paste0(day,"\n",wkdy)), family="Roboto Condensed", lineheight=0.8, size=4) + 
  labs(x="", y="", title="Another Type of Calendar") +
  scale_color_viridis_c()+
  theme(legend.position="bottom")

```





