---
title: Making Original Bingo - Heart Theme
author: Chisato
date: '2019-01-06'
slug: making-original-bingo-card-heart-theme
categories:
  - R
tags:
  - tidyverse
  - rvest
  - patchwork
description: ''
featured: ''
featuredalt: ''
featuredpath: ''
linktitle: ''
---

I have learned how to draw [a heart with mathmatical equation](http://mathworld.wolfram.com/HeartCurve.html), in fact there are so many "curves" you can draw with equation. Amazing, right?!?  You can find all sorts of different curves on [Wolfram Mathworld site](http://mathworld.wolfram.com/topics/PolarCurves.html). I'm really curious how did people find equation itself for [some of shapes](https://www.wolframalpha.com/input/?i=Obama+curve)? 

Also at last family reunion, I've played "[Bingo](https://en.wikipedia.org/wiki/Bingo_(U.S.))", which I haven't played for ages!  It's a great game when you have wide range of age groups, because kids were having tons of fun, but so were adults and grand parents! 


So I've decided to create little twist on bingo...! 
Instead of just drawing numbers between 1-75, you can play bingo with words, pictures, since I just really wanted to use my new "heart shape drawing shape..."  I decided I'll make up bingo with some words related to "Valentine's day".  


## Drawing Heart Shape with ggplot2

You can look at past entry on how to draw [cannabis](https://chichacha.netlify.com/2018/10/17/cannibus-curve-with-ggplot2/) here.  Same concept.  

There were 6 different heart shape example on Wolfram site, but I liked 6th heart the best for shape. You can look at equation [here](http://mathworld.wolfram.com/HeartCurve.html)


```{r dataframe_generator, echo=TRUE, fig.height=6, fig.width=11, message=FALSE, warning=FALSE, paged.print=FALSE}

library(tidyverse)
library(patchwork)

##  Function to generate heart shape around point xc and rc with some sizing.  Output of function is data frame with bunch of points which you can use to draw a heart!
# http://mathworld.wolfram.com/HeartCurve.html
heart_vertices <- function(xc,yc,size,npoints=100,...){
  #area = pi*r*r for circle... this heart has area of 180 = r*r
  radius = size*0.05  ## I'm not set on this...  I just wanted to make size smaller 
  yc = yc + 0.1*radius  ## Just adjusting center of heart bit
  t = seq(-pi,pi, length.out=npoints+1)
  x = xc + 16 * radius *(sin(t))^3
  y = yc + radius*13*cos(t) - radius*5*cos(2*t) - radius*2*cos(3*t) - radius*cos(4*t)
  df <- tibble(theta=t,x=x,y=y) 
  return(df)
}

## Above function with generate points you'd need to generate heart. If you adjust npoints to be higher, then you can draw smoother shape.

heart_vertices(0,0,1) %>%
  ggplot() +
  geom_line(aes(x=theta, y=x), color="#D22042de") +  
  geom_line(aes(x=theta, y= -x),color="#D22042de",linetype=3) +  ## this is just to make a design
  geom_line(aes(x=theta, y=y), color="#30C4C9de") +
  geom_line(aes(x=theta, y= -y), color="#30C4C9de", linetype=3) + ## this is just to make a design
  geom_polygon(aes(x=x,y=y), fill="#615375de", ## to draw heart use x=x,y=y
               color="#61537520",size=10) + ## I'm just playing around with line around filled part. 
  theme_minimal(base_family="Roboto Condensed") +
  scale_x_continuous(breaks=c(-pi,-pi/2,0,pi/2,pi), labels=c("-pi","","0","","pi")) +
  coord_fixed() +
  labs(x="",y="", caption="pink solid line = x value & blue solid line = y value")


```


## Drawing Bunch of Hearts on Grid

Now I know how to draw a single heart, I want to be able to plot bunch of hearts on a grid.  Since function heart_vertices creates data frame for a single heart around point xc and yc, I can create a grid with coordinates xc and yc.

```{r making_grid, echo=TRUE, fig.height=5, fig.width=11, message=FALSE, warning=FALSE, paged.print=FALSE}

my_grid <- tibble(
  xc=rep(c(1:15), times=5), 
  yc=rep(c(1:5), each=15),
  size=0.6, 
  id = c(1:75)
) 

my_grid %>% ggplot(aes(x=xc,y=yc)) +
  geom_point(shape=21, size=10) +
  geom_text(aes(label=id), family="Roboto Condensed", fontface="bold") +
  theme_minimal() +
  coord_fixed()

## For each points on grid generate points to draw heart 
print_out_base <-my_grid %>% 
  mutate(v=pmap(.,heart_vertices)) %>%
  unnest(v) %>% 
  ggplot(aes(x=x,y=y,group=id)) +
  geom_polygon(aes(fill=factor(yc))) +
  coord_fixed() +
  theme_minimal(base_family="Roboto Condensed") +
  scale_x_continuous(breaks=seq(0.5,16.5, by=1), labels=NULL) +
  scale_y_continuous(breaks=seq(0.5,5.5,by=1), labels=NULL) +
  scale_fill_manual(values=c("#30bcad","#57a337","#f8b620","#e03426","#eb73b3"), guide="none")

print_out_base

```

## Putting Words On Top of Hearts

Now I have the hearts placed on grid, I want some words on top. So I took inspiration from candies with sayings that I often see during Valentine's day week, which I didn't know the name of candy, but I think it's called  [Necco Sweetheart](https://en.wikipedia.org/wiki/Necco). 


```{r draw_cards, echo=TRUE, fig.height=7, fig.width=12, message=FALSE, warning=FALSE, paged.print=FALSE}

## Needs at least 75 words.... since there are 75 heart with some word placed on it.
## Some are not from those candies, I just made some up.  
love_msgs <- tibble(
  msgs = c("143", "#1 FAN", "#LOVE", "1-800\nCUPID", "HUG ME", 
            "KISSES", "BE MINE", "CRAZY\n4 U", 
            "HOLD\nHANDS", "UR\nLOVED", "PURR FECT", "WOO",
            "QT PIE", "RECIPE\n4 LOVE", "RISING STAR", "TABLE\nFOR 2", 
            "TOO SWEET", "TWEET", "TWO HEARTS", "TXT ME", "UR HOT", 
            "WHATS UP", "DESTINY", "WICKED COOL", "WINK\nWINK", 
            "STUNNING", "XOXO", "YOU&ME", "YUM\nYUM","SOUL\nMATE","BABE",
            "SAY YES","HELLO","DREAM\nBIG","BFF","HIGH\nFIVE","AWESOME",
            "SMILE","UR\nGR8","PHONE\nME","LOVE\nBIRD","BE TRUE","SURE LOVE",
            "MY BABY","HI GORGEOUS","HOT\nSTUFF","ADORE\nME","FUN","LOL","CALL ME",
            "PICK ME","DEAR\nONE","EVER\nAFTER","LOVER","ALL\nMINE","ANGEL",
            "RU SHY","SWEET PEA","LOVE\nBUG","ADORABLE","EMBRACE","FLOWERS",
            "CHERISH","CHOCOLATE","CUPCAKES","CRUSH","SECRET\nADMIRER",
            "VALENTINE","DOVES","LOVEBIRDS","DIAMONDS","PAARTY","HONEY",
            "PASSION","AWWW")
)

## Here you want to make sure you have at leat 75 words!    
love_msgs <- love_msgs %>% 
  arrange(msgs) %>% ## sort them alphabetically.... It makes it easier to find word that were picked out.
  mutate(idx=row_number()-1,
         row_group=floor(idx/15)+1) 


print_out_base +
  geom_text(aes(x=xc,y=yc, 
                label=love_msgs$msgs), 
            color="#ffffffde",family="Roboto Condensed", size=3, fontface="bold",
            lineheight=0.8,
            data=. %>% filter(theta==pi)) +
  labs(title="Print & Cut Them Into Pieces & Draw Them Out of Hat or Box",x="",y="", 
       caption="")

```

## Making Bingo Cards 

Similariy, now I want to make 5 x 5 bingo cards that each person gets to participate in the game. I've generated 4 cards as example. 

```{r bingo_playing_cards, echo=TRUE, fig.height=12, fig.width=12, message=FALSE, warning=FALSE, paged.print=FALSE}

## Making Bingo Cards (Base Design)
bingo_base <-tibble(
  xc = rep(c(1:5),times=5),
  yc = rep(c(1:5),each=5),
  size=0.6,
  id = c(1:25)
) %>% mutate(v=pmap(., heart_vertices)) %>%
  unnest(v) %>%
  ggplot(aes(x=x,y=y,group=id)) +
  geom_polygon(aes(fill=factor(xc))) +
  geom_polygon(fill="#000000de", data=. %>% filter(xc==3,yc==3))+
  theme_minimal(base_family="Roboto Condensed") +
  scale_x_continuous(breaks=c(1,2,3,4,5),labels=c("B","I","N","G","O"), position="top") +
  scale_y_continuous(labels=NULL) +
  scale_fill_manual(values=c("#30bcad","#57a337","#f8b620","#e03426","#eb73b3"), guide="none") +
  labs(x="",y="") +
  coord_fixed()

## Just to make card little more fun, let's add some quotes about love on each cards.
library(rvest)
love_quotes <- read_html("https://lifehacks.io/inspirational-love-quotes-sayings/") %>%
  html_nodes("h2") %>% html_text()
love_quotes <- love_quotes[2:64]

## Creating function to create one bingo card with randomly selected words on each rows.
make_card  <- function(name="") {
  love_msgs_list <- love_msgs %>% split(.$row_group)
  
  unique_card <- tibble(
    xc = rep(c(1:5),each=5),
    yc = rep(c(1:5),times=5),
    ## from each lists i want to select 5 randomly.
    msg = love_msgs_list %>% map(.,"msgs") %>% map(.,sample,5) %>% unlist()
  )
  unique_card <- unique_card %>% mutate(msg=ifelse(xc==3&yc==3,"FREE",msg))
  
  bingo_card <- bingo_base +
    geom_text(data=unique_card, aes(x=xc,y=yc,label=msg, group=NULL),
              family="Roboto Condensed", fontface="bold", color="white", size=3) +
    labs(title=str_c(name),caption=sample(love_quotes,size=1)) 
  
  bingo_card
  
}

## using patchwork, I want to print 4 cards
make_card("BINGO CARD 1") + make_card("BINGO CARD 2") + make_card("BINGO CARD 3") +
  make_card("BINGO CARD 4") + patchwork::plot_layout(ncol=2)


```

## Bonus : Drawing Flowers To Go With Hearts 

Just thought it would also be nice to draw flowers too.  After all, flowers go with hearts :). You can read more about rose [curve here](http://mathworld.wolfram.com/Rose.html) or [here](https://en.wikipedia.org/wiki/Rose_(mathematics))

```{r flower_drawing, echo=TRUE, fig.height=5, fig.width=12, message=FALSE, warning=FALSE, paged.print=FALSE}

flower_vertices <- function(xc,yc,radius,k=5,npoints=300,...){
  t = seq(0,2*pi, length.out=npoints+1) 
  m = sqrt(radius) * cos(k * t)
  x = xc + m * cos(t)
  y = yc + m * sin(t)
  df <- tibble(t=t,x=x,y=y,r=m)
  return(df)
}

flower_vertices(0,0,1,7) %>% 
  ggplot(aes(x=t)) +
  geom_line(aes(y=x), color="red", linetype=3) +
  geom_line(aes(y=y), color="blue",linetype=3) +
  geom_polygon(aes(x=x,y=y), alpha=0.5) +
  theme_minimal(base_family="Roboto Condensed") +
  coord_fixed() +
  labs(title="Rose Curve with K=7 - Flower with 7 Petals")


```

```{r grid_of_flowers, echo=TRUE, fig.height=12, fig.width=12, message=FALSE, warning=FALSE, paged.print=FALSE}

tibble(
  xc=rep(c(1:5),time=5),
  yc=rep(c(1:5),each=5),
  radius=0.1,
  k = c(1:25),
  id=c(1:25)
) %>% mutate(v=pmap(.,flower_vertices)) %>%
  unnest(v) %>%
  ggplot(aes(x=x,y=y,group=id)) +
  geom_polygon(aes(fill=id%%2)) +
  geom_point(aes(x=xc,y=yc), data=. %>% count(id,xc,yc), size=3,shape=19, alpha=0.7) +
  geom_text(aes(x=x,y=y, label=k), family="Roboto Condensed",
            size=8, vjust=1, fontface="bold", color="#000000ae",
            data=. %>% group_by(id) %>% filter(max(t)==t)) +
  theme_void(base_family="Roboto Condensed") +
  coord_fixed() +
  scale_y_reverse() +
  scale_fill_viridis_c(begin=0.2,end=0.7,option="magma", guide="none", alpha=0.8) +
  labs(title="Rose Curves with differnt k ",
       subtitle="r = cos(k * theta) ") 


```

## Flower Needs Butterfly Too..

There's also another called "[butterfly curve](http://mathworld.wolfram.com/ButterflyCurve.html)".  

I think flower deserves butterfly...  So here's butterfly curve drawn in similar manner as the above. 

```{r butterfly, echo=TRUE, fig.height=7, fig.width=12, message=FALSE, warning=FALSE, paged.print=FALSE}


butterfly_vertices <- function(xc,yc,npoints=1000,...){
  t = seq(0,12*pi, length.out=npoints+1)
  x = xc + sin(t)*(exp(cos(t))-2*cos(4*t)-sin(t/12)^5)
  y = yc + cos(t)*(exp(cos(t))-2*cos(4*t)-sin(t/12)^5)
  df <- tibble(x=x,y=y,t=t) %>% mutate(pos=row_number())
  return(df)
}

ggplot() +
  geom_path(data=butterfly_vertices(1,1),aes(x=x,y=y, color=pos)) +
  geom_polygon(data=butterfly_vertices(8,1), 
               aes(x=x,y=y,fill=factor(floor(t/pi))), color="#000000de") +
  coord_fixed() +
  theme_void() +
  scale_fill_viridis_d(alpha=0.3, guide="none") +
  scale_color_viridis_c(option="magma", guide="none")


```

